---
title: "DS3001 Final Project"
author: "Joseph Lee (sl5nj), Iain Muir (iam9ez), Kent Williams (kbw3bb)"
date: "12/6/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DATA PREPARATION 
## Step 0 — Import Libraries
randomForest, rio, plyr, tidyverse, rpart, psych, pROC, rpart.plot, rattle, caret, mlbench, MLmetrics, ROCR, mltools, data.table, plotly
```{r, include=FALSE}
library(randomForest)
library(rio)
library(plyr)
library(tidyverse)
library(rpart)
library(psych)
library(pROC)
library(rpart.plot)
library(rattle)
library(caret)
library(mlbench)
library(MLmetrics)
library(ROCR)
library(mltools)
library(data.table)
library(plotly)
```

## Step 1 — Load Data
Data from Kaggle... Summary/Description
```{r include=FALSE}
data <- read_csv('/Users/iainmuir/Desktop/4Y 1S/DS 3001/Practice-of-Data-Science/Final Project/data.csv')
dass = data
```

```{r echo=FALSE}
head(dass)
```

```{r echo=FALSE}
dim(dass)
```

```{r include=FALSE}
column_index <- tibble(colnames(dass))
column_index
```

## Step 2 — Handle Missing Values

Define Empty Strings cells as NA
```{r include=FALSE}
dass[dass == ""] <- NA
```

```{r include=FALSE}
na_count <-sapply(dass, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
na_count
```

```{r echo=FALSE}
cat("Total Missing Values:", sum(na_count$na_count))
```

Remove "major" variables; ~ 1/3 of data points are NULL
```{r include=FALSE}
dass <- dass[, -172]
dim(dass)
```

```{r include=FALSE}
na_count <-sapply(dass, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
```

```{r echo=FALSE}
cat("Total Missing Values:", sum(na_count$na_count))
```

```{r include=FALSE}
dass <- dass[complete.cases(dass),]
```

```{r echo=FALSE}
dim(dass)
```

## Step 3 — Ensure Correct Data Types
```{r include=FALSE}
dass[,158:171] <- lapply(dass[,158:171], as.factor)
```

```{r include=FALSE}
str(dass)
```

## Step 4 — Collapse Factor Variables
Explore Distributions of Factor Variables
```{r include=FALSE}
table(dass$country)     # Malaysia, USA, Other
table(dass$education)   # FINE --> Remove 0s?
table(dass$urban)       # FINE --> Remove 0s?
table(dass$gender)      # FINE --> Remove 0s?
table(dass$engnat)      # FINE --> Remove 0s?
table(dass$religion)    # Muslim, Christian, Other
table(dass$orientation) # FINE --> Remove 0s?
table(dass$race)        # Asian, White, Other
table(dass$voted)       # FINE --> Remove 0s?
table(dass$married)     # FINE --> Remove 0s?
```

Factor Collapse
* country: Malaysia, USA, Other
* religion: Muslim, Christian, Other
* race: Asian, White, Other
```{r include=FALSE}
dass$country <- fct_collapse(
  dass$country,
  Malaysia="MY",
  USA="US",
  other_level="Other"
)
dass$religion <- fct_collapse(
  dass$religion,
  Muslim="10",
  Christian=c("4", "5", "6", "7"),
  other_level="Other"
)
dass$race <- fct_collapse(
  dass$race,
  Asian="10",
  White="60",
  other_level="Other"
)
```

## Step 5 — Split Data into Train/Tune/Test Sets

```{r include=FALSE}

```

# UNSUPERVISED MACHINE LEARNING
## K-Means Clustering

```{r include=FALSE}

```

# SUPERVISED MACHINE LEARNING

## Step X — Manually Create Target Variable
```{r include=FALSE}
col_vect <- pull(column_index, `colnames(dass)`)
col_vect
```

```{r include=FALSE}
depression <- c(3, 5, 10, 13, 16, 17, 21, 24, 26, 31, 34, 37, 38, 42)
anxiety <- c(2, 4, 7, 9, 15, 19, 20, 23, 25, 28, 30, 36, 40, 41)
stress <- c(1, 6, 8, 11, 12, 14, 18, 22, 27, 29, 32, 33, 35, 39)
d <- ifelse(substr(col_vect, 2, nchar(col_vect) - 1) %in% depression, col_vect, NA)
d <- d[!is.na(d)]
a <- ifelse(substr(col_vect, 2, nchar(col_vect) - 1) %in% anxiety, col_vect, NA)
a <- a[!is.na(a)]
s <- ifelse(substr(col_vect, 2, nchar(col_vect) - 1) %in% stress, col_vect, NA)
s <- s[!is.na(s)]
```


## Logistic Regresion
```{r include=FALSE}

```

## Random Forest
```{r include=FALSE}

```
